---
title       : Analysis of Salaries of Professors
subtitle    : 2008-09 Faculty Salary Survey from Salaries {car}
author      : K White
job         : Coursera/Johns Hopkins Data Science Specialization Student
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Data Background

* The data source collected nine-month academic salaries for three faculty levels at a US college in 2008-09 as part of an on-going monitoring program run by the college administration to follow salary differences between male and female faculty members  
  
* The data had 397 observations tracking faculty rank, departmental discipline type, years since PhD received, years of service, gender, and salary

* This application invites exploratory data discovery by breaking out three elements of the faculty salaries: faculty rank, gender, and departmental discipline type

* All three breakdowns have a point plot comparing the categories within the variables color coded for Gender; an optional trend line can be displayed as can a rug frequency gauge on each axis

* Additionally, summary statistics are calculated for each breakdown, providing the category's salary mean, median, maximum and minimum values

--- 

## Data Adjustments

The data was adjusted from its raw format by renaming relevant columns to improve presentation quality and replacing abbreviated values in columns with longer or descriptive values. No other adjustments were performed on the data set.
```{r setoptions, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
library(car)
library(dplyr)
data(Salaries)
selectData <- Salaries
selectData <- rename(selectData, 
                     Gender = sex, 
                     Salary = salary, 
                     Rank = rank, 
                     Discipline = discipline)
selectData$Discipline <- sapply(selectData$Discipline, function(y) as.character(gsub("^A$", "Theoretical Discipline Department", y)))
selectData$Discipline <- sapply(selectData$Discipline, function(y) as.character(gsub("^B$", "Applied Discipline Department", y)))
selectData$Rank <- sapply(selectData$Rank, function(y) as.character(gsub("^Prof$", "Professor", y)))
selectData$Rank <- sapply(selectData$Rank, function(y) as.character(gsub("^AsstProf$", "Assistant Professor", y)))
selectData$Rank <- sapply(selectData$Rank, function(y) as.character(gsub("^AssocProf$", "Associate Professor", y)))
```
```{r headdata, echo=FALSE, warning=FALSE, message=FALSE}
head(selectData)
```

---

## Data Summary Exploration
```{r postvalues, echo=FALSE, warning=FALSE}
# head(selectData)
summary(selectData)
```

---

## Areas for Further Exploration

The trendline and statistics tables provide intriguing insights that may warrant further inquiry:
* There are far fewer females employed than males
* The max tenure of females is only slightly over half that of males
* On average, females trail male counterparts in salary by $15,000 but have a higher minimum salary by about $5,000
* Applied disciplinary departments have a better compensation expectation than theoretical disciplinary departments after 20 years of service


